Methods for Large Scale Learning to Rank

### 2014-09-16 report/Chapters/Implementation.tex
uitbreiding implementation

\section{Architecture}
Ranking algorithms consist of a sequence of operations on the input data which are often, but not always, of iterative nature. Apache Pig \cite{Olston2008} will be used to implement the sequence of operation on the input data, because the higher level of abstraction that Apache Pig offers makes it easier to focus solely on the data operations as compared to native Hadoop MapReduce.\\

\subsection{Storing Input Data for HDInsight Processing}
Azure HDInsight supports both the traditional \ac{HDFS} as described in \cite{Shvachko2010}, as well as Microsoft's own \ac{WASB} storage solution. Blob storage decouples the storage from the HDInsight Hadoop cluster; it enables safe deletion of a HDInsight cluster without losing data, as data is not solely stored on the cluster itself, but also on a separate storage that is not cluster-dependent. Azure \ac{WASB} storage allows the user to select one of their data centre for storage. \ac{WASB} storage in the West Europe region (located in Amsterdam) will be used for storage, as this data centre is located close to where the experiments are executed.

@@

% TODO: Stuk over werkzaamheden framework uitbreiden: uitlezen resultaten (evaluatieresultaten + doorgeven tussenresultaten (die tweede wellicht later in het document pas introduceren))
In a cross-validation evaluation experiment of a ranking model, the handling of which folds are loaded for training, evaluation and testing is a task that need to be taken care of regardless of the ranking model. Given that most ranking models are of an iterative nature, the task of iteration handling is also a task that need to be done for (most) models. Since these tasks need to be considered for (almost) all ranking models it seems natural to develop a framework that takes care of these tasks. The aim of this framework is to let implementations of ranking models focus solely on implementing the sequential steps of one iteration of the algorithm, while the framework arranges that 1) these sequential steps are performed iteratively and 2) these sequential steps are performed on the multiple training folds of data. This framework that handles folding and iterations will be implemented in Java and will work such that fold- and iteration dependent parts of the Pig code will be generated dynamically by the Java framework after which the Pig job will be sent to the cluster.\\

>>>>

\section{Architecture}
Ranking algorithms consist of a sequence of operations on the input data which are often, but not always, of iterative nature. Apache Pig \cite{Olston2008} will be used to implement the sequence of operation on the input data. Pig Latin, the data processing language that runs on Apache Pig, designed based on the observation that the traditional MapReduce paradigm is too low-level and rigid, and holds the middle between the declarative style of SQL and the procedural style of MapReduce. The Apache Pig system translates Pig Latin into MapReduce plans that are executed over Hadoop. The choice to implement the Learning to Rank algorithms in Pig Latin allow more focus on the data operations and less focus on low-level implementation details compared to native Hadoop MapReduce. Furthermore, it allows us to rely on Apache Pig to create efficient MapReduce plans out of the Pig Latin code which lowers the implementation-dependent factor of the experiments.\\

\subsection{Storing Input Data for HDInsight Processing}
Azure HDInsight supports both the traditional \ac{HDFS} as described by Shvacko et al. \cite{Shvachko2010}, as well as Microsoft's own \ac{WASB} storage solution. Blob storage decouples the storage from the HDInsight Hadoop cluster; it enables safe deletion of a HDInsight cluster without losing data, as data is not solely stored on the cluster itself, but also on a separate storage that is not cluster-dependent. Azure \ac{WASB} storage allows the user to select one of their data centre for storage. \ac{WASB} storage in the West Europe region (located in Amsterdam) will be used for storage, as this data centre is located close to where the experiments are executed.

@@

% TODO: Stuk over werkzaamheden framework uitbreiden: uitlezen resultaten (evaluatieresultaten + doorgeven tussenresultaten (die tweede wellicht later in het document pas introduceren))
Fold handling  in a cross-validation experiment is the process of loading the correct data folds for training, validation and testing in the multiple rounds that the cross-validation experiment consists of. Fold handling is a task that needs to be taken care of regardless of the ranking model that is being evaluated. Given that most ranking models are of an iterative nature, the task of iteration handling is also a task that need to be done for most models. Iteration handling and fold handling are procedures in which the same steps are repeated for each iteration or cross-validation round respectively. Pig Latin, however, has no support for loops which are needed to perform iteration handling and fold handling. Since iteration handling and fold handling are tasks that need to be addressed for each ranking model and that cannot be solved with Pig, we create a framework that takes care of both iteration and fold handling and generates the Pig Latin code for the current iteration and fold. The aim of this framework is to let implementations of ranking models focus solely on implementing the sequential steps of one iteration of the algorithm, while the framework arranges that 1) these sequential steps are performed iteratively and 2) these sequential steps are performed on the multiple training folds of data. This framework that handles folding and iterations will be implemented in Java and will work such that fold- and iteration dependent parts of the Pig code will be generated dynamically by the Java framework after which the Pig job will be sent to the cluster.\\

Another responsibility of the framework is to handle communication between multiple Pig jobs within a single iteration of a Learning to Rank algorithm. The framework enables reading the result of a Pig job from \ac{WASB} storage which can then be used as parameter within a subsequent Pig job. An example of this is a Pig implementation of gradient descent, where a first Pig job might calculate gradients and writes them to storage, after which the framework assists in reading the gradients from storage and allows them to be used as input parameters of a second Pig job that calculates new feature weights based on its gradient parameter and a predetermined step size.\\
